---
title: Report on a dataset in the AusFizz compilation
output:
  html_document:
    df_print: kable
    highlight: tango
    keep_md: no
    smart: no
    theme: yeti
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
params:
    dataset_id: provide
    database: provide
editor_options:
  chunk_output_type: console
---

<!-- hack to get indentation on 3rd level of floating TOC; see
https://stackoverflow.com/questions/46201753/rmarkdown-indentation-of-toc-items-in-html-output
 -->

<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
# knitr defaults
knitr::opts_chunk$set(echo = FALSE, cache = FALSE)

# default for table format
options(knitr.table.format = "html")

# remove warnings from dplyr about "summarise"
options(dplyr.summarise.inform = FALSE)

# Guidelines for writing report code
# - use tidyverse style and format: http://htmlpreview.github.io/?https://github.com/nicercode/2018_BEES_regression/blob/master/tidyverse.html
# - use kableExtra for styling: https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
# - use knitr chunk options: https://rmarkdown.rstudio.com/lesson-3.html

# Some useful functions


#' Generate hyperlink for markdown and html
#'
#' Generate hyperlink for markdown and html files
#'
#' @param link character string for the URL link
#' @param text character string for the text to display
#' @param type file type, default is markdown "md" otherwise html
#'
#' @return character string with the text and link formatted for md and html
#' @export
#'
#' @examples as_link("www.austraits.org", "austraits")
as_link <- function(link, text, type = "md") {
  if (type == "md") {
    sprintf("[%s](%s)", text, link)
  } else {
    sprintf("<a href='%s'> %s </a>", link, text)
  }
}


#' Get SHA link from GitHub
#'
#' Get SHA link using the util_get_SHA() function. The link generated leads to the latest
#' commit for the GitHub repository. SHA is the abbreviated SHA-1 40-digit
#' hexadecimal number which GitHub uses to track commits and changes made to a repository.
#'
#' @param ... arguments passed to the util_get_SHA()
#'
#' @return SHA link to a GitHub commit as a character string formatted using markdown syntax
util_get_SHA_link <- function(...) {
  sha <- util_get_SHA(...)
  as_link(sprintf("https://github.com/traitecoevo/responses.build/tree/%s", sha), sha)
}

```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
library(responses.build)
library(dplyr)
library(knitr)
library(kableExtra)

`%||%` <- function(a, b) if (is.null(a) || (length(a) == 1 && is.na(a))) b else a

my_kable_styling <- util_kable_styling_html

# `definitions` is the vocabulary of variables -- the readings taken at each
# point of a curve. There is no second vocabulary of derived parameters:
# fitting curves is out of scope for this compilation.
definitions <- database$definitions
data_types <- database$data_types %||% list()

data_study <- extract_dataset(database, dataset_id)
metadata <- read_metadata(file.path("data", dataset_id, "metadata.yml"))
schema <- database$schema

#' Render a ggplot to a PNG and embed it at a chosen size
#'
#' knitr's chunk-level `fig.height` applies uniformly to every plot in a chunk,
#' so a per-plot size cannot be set that way inside a loop.
embed_plot <- function(plot, height = 4, width = 8) {
  file <- tempfile(fileext = ".png")
  ggplot2::ggsave(file, plot, width = width, height = height, dpi = 150)
  sprintf("<img src='data:image/png;base64,%s' style='max-width:100%%'>",
          base64enc::base64encode(file))
}

# Data release (not code/package release) of APCalign used to align taxon
# names -- recorded as a dated GitHub release tag among the compiled
# database's `related_identifiers` metadata, e.g.
# ".../APCalign/releases/tag/2025-04-15".
apcalign_identifier <- database$metadata$related_identifiers %>%
  purrr::detect(~ !is.null(.x$identifier) && grepl("APCalign", .x$identifier, ignore.case = TRUE))

apcalign_data_release <- if (!is.null(apcalign_identifier)) {
  sub(".*/tag/", "", apcalign_identifier$identifier)
} else {
  "unknown"
}

# start notetaker device
questions <-  responses.build:::notetaker_start()

new_question <- function(txt) {
  txt_formatted <- sprintf("<font color='orange'>**Question:** %s</font>", txt)
  x <- questions(responses.build:::notetaker_add_note, responses.build:::notetaker_as_note(txt_formatted)) %>%
    responses.build:::notetaker_print_notes(as_anchor = TRUE)
  writeLines(sprintf("%s\n\n", x))
}

new_question_2 <- function(txt) {
  txt_formatted <- sprintf("<font color='blue'>**Dataset-specific question:** %s</font>", txt)
  x <- questions(responses.build:::notetaker_add_note, responses.build:::notetaker_as_note(txt_formatted)) %>%
    responses.build:::notetaker_print_notes(as_anchor = TRUE)
  writeLines(sprintf("%s\n\n", x))
}

# Variable to record when definitions have been specified

structures_already_defined <- c("names")


```

# Introduction

`AusFizz` is an open-source compilation of plant physiological response-curve data for Australian
flora -- leaf gas exchange, hydraulic vulnerability, and fluorescence. Where most trait compilations
hold one value per plant, this one holds **curves**: sets of readings taken across a driver gradient,
whose points belong together and only mean something together.

That distinction runs through the whole database. A single A--Ci curve is not a hundred separate
measurements; it is one measurement, of a response. AusFizz keeps it that way, so that the curve can
be retrieved, plotted and fitted as the object it is.

Three kinds of thing are recorded, and they are not interchangeable:

| | | |
|---|---|---|
| **variables** | the readings taken at every point of a curve | `A`, `Ci`, `gsw`, `Tleaf`, `PSIstem` |
| **curves** | the curve itself, as something you can ask for | one A--Ci curve, one vulnerability curve |
| **traits** | parameters fitted from a curve, one value per plant | `Amax`, `Vcmax25`, `P50`, `TLP` |

The curves are the point of this compilation; the traits fitted from them are what eventually
reaches [AusTraits](https://austraits.org), the wider compilation of Australian plant traits.

Data are assembled by [responses.build](https://github.com/traitecoevo/responses.build), which keeps
all the metadata that came with a measurement -- methods, instrument, location, experimental
treatment -- attached to it. Plant names are aligned to the
[Australian Plant Census](https://biodiversity.org.au/nsl/services/search/taxonomy) using
[APCalign](http://traitecoevo.github.io/APCalign).

AusFizz exists because researchers contribute to it. Thank you for being part of making
physiological data for Australia's flora reusable.

## Purpose of this report

This document describes the information we have for the study **` `r dataset_id` `**. 

This report gives both you, as a data contributor, and us, as AusFizz compilers, the opportunity to:

1. Verify that our understanding and handling of your data is correct
2. Provide any missing information about the data

Each study is assigned a unique `dataset_id`, with information for that dataset in its own folder, organised into two files:

- ` `r dataset_id`/data.csv`: the readings themselves, as they came from the instrument
- ` `r dataset_id`/metadata.yml`: everything about them -- which instrument, which methods, which treatment, and how the columns map onto AusFizz variables.

**The section that most repays your attention is [The curves](#the-curves)**, where every curve in
the dataset is drawn. A curve that looks wrong there is wrong, whatever the summary tables say.

Data from those files have been used to generate the report below.

Please review the report, check it is correct, and send updates where appropriate. 

There are two sets of queries in this report:

1. <font color='orange'>*Queries in orange scattered throughout the report.*</font> These are automatically generated for all reports and do not indicate we have a concern about your dataset submission; consider them prompts to draw your attention to each section of the report.

2. <font color='blue'>*Queries in blue that (might) appear at the start of the report.*</font> These are explicit questions pertaining to your dataset that the dataset curator had when adding your study.

All questions appear again as a list at the end of the document.

To aid your review, we have attached the following files:

- `r dataset_id`/data.csv and `r dataset_id`/metadata.yml: The raw inputs for your study.

### Send feedback by

1. **Copying and pasting the questions at the end of the report into an email and answering as many as possible.**

   For the automated (orange) questions, if the pertinent metadata or data plot/table match your expectations, simply answer "yes".
   
   The targeted (blue) questions represented an uncertainty or concern the AusFizz data curator had as they merged in your study. Please consider these carefully, as they usually indicate missing metadata or some source of ambiguity (e.g. units not provided, uncertainty about trait value substitutions, study site latitude/longitude).
    
2. Sending back your responses and any additional metadata required.

3. If appropriate, send a new spreadsheet to replace the existing `data.csv` file.


```{r, results='asis', echo=FALSE}
# Any additional questions from metadata

if (!is.null(metadata$questions) && !is.na(metadata$questions[1])) {

  x <-
    names(metadata$questions)[!(names(metadata$questions) %in% c("additional_traits", "austraits"))]

  if (length(x) > 0)
    writeLines("# Major questions\n\n")

  for (i in seq_along(x))
    new_question_2(sprintf("(section `general`) %s", metadata$questions[[i]]))
}
```


# Study details

We document information about the context of the trait measurements, including both dataset-specific metadata and metadata pertaining to individual trait measurements. Dataset-level metadata includes details on the people involved, the sources, and the overall sampling strategy. Observation-level metadata includes trait measurement methods, locations, location properties, collection date, etc.

## Contributors

The following people are listed as data collectors in this study:

```{r, results='asis', echo=FALSE}
metadata$contributors$data_collectors %>%
  convert_list_to_df2() %>%
  my_kable_styling()
```

We also require an email address for all people. People involved in the study who cannot be tracked down need to be designated as assistants.

In addition, the following people are included as

- assistants: `r ifelse(is.null(metadata$contributors$assistants), "none", paste(metadata$contributors$assistants, collapse=", "))`
- data curator(s):  `r metadata$contributors$dataset_curators`

```{r, results='asis', echo=FALSE}
new_question("(section `people`) Are all appropriate people listed, with appropriate details?")
```

## Sources

Data from this study are recorded as coming from the following sources:

Primary citation:

```{r, results='asis', echo=FALSE}
metadata$source$primary %>%
  convert_list_to_df1() %>%
  my_kable_styling()
```

```{r, results='asis', echo=FALSE}
n <- length(metadata$source)
if (n > 1) {
  writeLines("Secondary citations:")

  for (i in 2:n) {
    metadata$source[[i]] %>%
    convert_list_to_df1() %>%
    my_kable_styling() %>%
    cat()
    writeLines("\n\n")
  }
}
```

```{r, results='asis', echo=FALSE}
new_question("(section `source`) Are the citation details for this study correct?")
```

## General description

The dataset-level metadata we have mapped in is as follows:

- the fields `taxon_name`, `trait_name`, `value` (for long-format datasets), & `location_name` indicate the column from `data.csv` we have mapped in for this information (the built table calls `trait_name` `variable`)
- `description` & `sampling_strategy` are based on information in the publication or notes you've provided
- `basis_of_record`, `life_stage`, and `collection_date` can be a fixed value for the dataset or link to a column that varies by observation
- `custom_R_code` is just that; code specifically written to manipulate `data.csv` as it is read into the traits.build workflow

```{r, results='asis', echo=FALSE}
metadata$dataset %>%
  convert_list_to_df1() %>%
  my_kable_styling()
```

```{r, results='asis', echo=FALSE}
new_question("(section `dataset`) Can you provide more detailed information for any of these variables?")
```

```{r, results='asis', echo=FALSE}
missing <- metadata$dataset %>%
  convert_list_to_df1() %>%
  filter(value == "unknown") %>%
  pull(key)

for (v in missing) {
  sprintf("(section `dataset`) Can you provide missing details for the variable `%s`?\n", v) %>%
    new_question()
  sprintf("The variable `%s` captures *'%s'*\n", v, schema$metadata$elements$dataset$values[[v]]) %>%
    writeLines()
}
```

## Locations

```{r, results='asis', echo=FALSE}
# When every record is species-level (or a higher taxonomic rank), 
# jump straight to a single confirmatory question instead.
is_species_only_dataset <- nrow(data_study$measurements) > 0 && isTRUE(all(data_study$measurements$entity_type == "species"))

if (is_species_only_dataset) {
  writeLines("This dataset only includes species-level trait observations. Such observations are not linked to a specific location or location properties.")

  writeLines("")

  new_question("(section `locations`) Is it correct that all your trait data apply to the species, not just specific populations or individuals, and therefore is it appropriate that there is no location data attached to these observations?")
}
```

```{r, results='asis', echo=FALSE, warning=FALSE}

create_if_missing <- function(df, var, type = rep(NA_character_, nrow(df))) {
  if (is.null(df[[var]]))
    df[[var]] <- type
  df
}

location_main <- c("location_name", "longitude_deg", "latitude_deg", "description")

if (!is_species_only_dataset) {
  writeLines("Data were collected at the following sites and locations. As a minimum, we are aiming to collect the variables `latitude`, `longitude`, and `description`.")

  # entity_type values recorded against a specific collection location.
  # Species-level (and higher taxonomic-rank) measurements are summary
  # statistics/expert scores for a taxon as a whole and never have a
  # location, so they're excluded here rather than showing up as a spurious
  # "unknown" location row full of NAs.
  location_entity_types <- c("individual", "population", "metapopulation")

  locations <-
    full_join(by = "location_id",
              # count of records by location
              data_study$measurements %>%
                filter(entity_type %in% location_entity_types) %>%
                # `ifelse()` degrades a character column to logical when the
                # data is empty (e.g. a dataset with no location-tied entity
                # types), which then fails to join against `location_id`
                # (character) below -- `coalesce()` doesn't have that problem.
                mutate(location_id = dplyr::coalesce(location_id, "unknown")) %>%
                group_by(location_id) %>%
                summarise(records = n()),
              # details on locations from metadata
              data_study$locations %>%
                filter(location_property %in% location_main) %>%
                select(-dataset_id) %>%
                spread(location_property, value) %>%
                create_if_missing("location_id") %>%
                create_if_missing("location_name") %>%
                create_if_missing("latitude_deg") %>%
                create_if_missing("longitude_deg") %>%
                create_if_missing("description")  %>%
                mutate(
                  latitude_deg = as.numeric(latitude_deg),
                  longitude_deg = as.numeric(longitude_deg)
                  )
    ) %>%
   select(location_id, location_name, records, latitude_deg, longitude_deg, description)

  if (nrow(locations) > 0) {
    locations %>%
      my_kable_styling() %>%
      writeLines()

    if (any(tolower(locations$location_name) %in% c(NA, "unknown"))) {
      new_question("(section `locations`) **Location data incomplete or unknown!** Can you provide location details where your data were sampled? If data are from a National Park, biological reserve, or other location with a name, but not linked to a specific location within the park, the location name should be that national park (or similar), and we will use the park's headquarters as the GPS coordinates. If the data are from herbarium specimens or a taxa monograph, it is appropriate to leave this information blank.")
    }
  } else {
    writeLines("There is currently no location data associated with this dataset, as all trait values are species-level (or higher taxonomic rank) and so are not tied to a specific collection location.")
  }
} else {
  # keep `locations` defined (0 rows) so the `maps` chunk below has
  # something safe to filter/check, without rendering anything itself
  locations <- tibble::tibble(
    location_id = character(), location_name = character(), records = integer(),
    latitude_deg = numeric(), longitude_deg = numeric(), description = character()
  )
}

```

```{r maps, eval= TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.keep="all"}

# Make a map of study locations using leaflet
locations <- locations %>% filter(!is.na(latitude_deg) & !is.na(longitude_deg))
if (nrow(locations) > 0) {
  leaflet::leaflet() %>%
  leaflet::addTiles() %>%
  leaflet::addScaleBar() %>%
  leaflet::addMiniMap(zoomLevelFixed = 2) %>%
  leaflet::addMarkers(lng = locations$longitude_deg, lat = locations$latitude_deg,
             label = locations$location_name)
}
```

```{r, results='asis', echo=FALSE}
if (!is_species_only_dataset) {
  new_question("(section `locations`) Do location details look complete and accurate? As a minimum we would like `latitude`, `longitude`, `description`.")
}
```

```{r, results='asis', echo=FALSE}
if (!is_species_only_dataset) {
  writeLines(c(
    "## Additional location data (optional)",
    "",
    "The following variables have been recorded for the locations included in the study:",
    ""
  ))

  locations_extra <- data_study$locations %>%
    filter(!(location_property %in% location_main)) %>%
    select(-dataset_id) %>%
    spread(location_property, value)

  if (nrow(locations_extra) > 0) {
    locations_extra %>%
      my_kable_styling() %>%
      writeLines()

    new_question("(section `locations`) Do the additional location details look complete?")
  } else {
    writeLines("There is currently no additional location data associated with this dataset.")
  }
}

```

## Contexts

```{r, results='asis', echo=FALSE}
# Create intro statement that rephrases information from traits.build schema
# definition into a more readable format. 
# This requires some word substitutions and syntax changes
category_options_text <- schema$metadata$elements$contexts$elements$category %>%
  stringr::str_remove("^[^.]*\\.\\s*") %>%
  stringr::str_replace("^Options", "options")


description_text <- schema$metadata$elements$contexts$description %>%
  stringr::str_remove("\\.\\s*$")
description_text <- paste0(tolower(substring(description_text, 1, 1)), substring(description_text, 2))

# Present what the 5 possible context property categories are
# since we report to the reader which category each of their context properties is in.
sprintf(
  "Contexts are %s. There are five context property categories: %s",
  description_text,
  category_options_text
) %>%
  writeLines()
```

That is, context properties are flexible and allow you to map in any additional metadata you have documented that might be useful for explaining differences in trait values. This ensures when your data are used in future analyses your expert knowledge about your trait measurements is documented. This might include plant age, measurement methods that subtely differed between trait measurements, fire history, or anything that can be added as a column of information to the `data.csv` file.


```{r, results='asis', echo=FALSE}
context_properties <- data_study$contexts %>%
  select(-dataset_id) %>%
  group_by(context_property) %>%
  group_split()

description_removed <- FALSE

if (length(context_properties) > 0) {
    writeLines("The following contexts have been recorded for the measurements included in this study.\n")

    # Helper function for properly spacing lines in output
    context_spacer <- function() {
      writeLines("")
      writeLines("<div style='margin-top: 2.0em;'></div>")
      writeLines("")
    }

    if ("replicate observations" %in% data_study$contexts$context_property) {
      writeLines("The context property `replicate observations` has been added because you have multiple trait observations at a population or species-level linked to the same value for `species` * `location` * `other contexts` * `value_type`, and our pivoting functions require an additional context property that flags each row as an independent observation.")
      context_spacer()
    }

    for (i in seq_along(context_properties)) {
      var_in <- context_properties[[i]][["var_in"]][1]
      category <- context_properties[[i]]$category[1]
      article <- ifelse(grepl("^[aeiouAEIOU]", category), "an", "a")
      context_property_name <- context_properties[[i]]$context_property[1]

      if (i > 1) {
        context_spacer()
      }

      writeLines(paste0("**", context_property_name, "** has been ",
        ifelse(is.null(var_in), "added ",
          paste0("mapped from the column titled **", var_in, "** in your dataset, ")),
        "as ", article, " `", category, "`."))

      writeLines("\n\nThe following values exist for this context:\n")

      context_table <- context_properties[[i]] %>% dplyr::select(dplyr::any_of(c("value", "description")))

      # As the compiled `contexts` table doesn't retain the originally supplied
      # values (those in `find` in the metadata.yml), 
      # these need to be pulled back in from metadata.yml for display
      metadata_context_entry <- metadata$contexts %>%
        purrr::detect(~ identical(.x$context_property, context_property_name))

      if (!is.null(metadata_context_entry) && !is.null(metadata_context_entry$values)) {
        find_value_table <- metadata_context_entry$values %>%
          convert_list_to_df2() %>%
          dplyr::select(dplyr::any_of(c("find", "value")))

        # A `values` block that lists only `value` -- no `find` -- has no
        # `find` column to select, and then `find` in the coalesce below
        # resolves to `base::find()`, which fails with "must be a vector, not a
        # function". Every context this build generates from a dataset-level
        # descriptor is of that shape, so add the column when it is missing.
        if (!"find" %in% names(find_value_table)) {
          find_value_table[["find"]] <- find_value_table[["value"]]
        }

        context_table <- context_table %>%
          dplyr::distinct() %>%
          dplyr::left_join(find_value_table, by = "value") %>%
          dplyr::mutate(find = dplyr::coalesce(.data$find, .data$value)) %>%
          dplyr::select(dplyr::any_of(c("find", "value", "description")))
      }

      if ("description" %in% names(context_table) && all(is.na(context_table$description))) {
        context_table <- context_table %>% dplyr::select(-description)
        description_removed <- TRUE
      }

      if (identical(names(context_table), "value")) {
        # If there are no contect value substitutions or descriptions
        # collapse the values into a single
        # sorted, comma-delimited, line-wrapping line.
        distinct_values <- context_table$value %>% unique()
        numeric_values <- suppressWarnings(as.numeric(distinct_values))
        sorted_values <- if (!anyNA(numeric_values)) {
          distinct_values[order(numeric_values)]
        } else {
          sort(distinct_values)
        }
        writeLines(" ")
        # helper function to add multiple scapes that would otherwise be collapsed
        value_separator <- paste0(";", strrep("&nbsp;", 3), " ")
        writeLines(paste0(sorted_values, collapse = value_separator))
        writeLines(" ")
      } else {
        writeLines(" ")
        context_table %>%
          my_kable_styling() %>%
          writeLines()
        writeLines("\n\n")
      }
    }
  new_question("(section `contexts`) Do the context details look complete?")
  if (description_removed) {
    new_question("(section `contexts`) For some contexts, we do not map in descriptions of individual context values; for instance, context values that are numbers, names of seasons, or species names. Let us know if you feel like omitting descriptions for any of your context values leads to ambiguity in their meaning.")
  }
  } else {
  writeLines("There is currently no contextual data associated with this dataset.")

  writeLines("")

  new_question("(section `contexts`) Are there context properties that should be included for this dataset?")
  }

```

## Identifiers

```{r, results='asis', echo=FALSE}
# Create intro statement that rephrases information from traits.build schema
# definition into a more readable format.

identifiers_desc_text <- schema$metadata$elements$identifiers$description %>%
  stringr::str_remove("^.*\\bidentifiers that\\s*") %>%
  stringr::str_remove("\\.\\s*$") %>%
  stringr::str_replace("^cross-references\\b", "cross-reference")

sprintf("Identifiers %s.", identifiers_desc_text) %>%
  writeLines()

writeLines("")

writeLines(
  ifelse(nrow(data_study$identifiers) == 0,
    "There are no identifiers provided that link data from this dataset to other datasets, biodiversity portals, collections, or other databases.",
    "The following identifiers are linked to the trait data observations for this study."))

writeLines("")

if (nrow(data_study$identifiers) > 0) {
  identifiers_meta <- metadata$identifiers %>% convert_list_to_df2()

  tibble::tibble(
    var_in = paste(unique(identifiers_meta$var_in), collapse = ", "),
    identifier_type = paste(unique(identifiers_meta$identifier_type), collapse = ", "),
    institution_code = paste(unique(identifiers_meta$institution_code), collapse = ", "),
    records = nrow(data_study$identifiers)
  ) %>%
    my_kable_styling() %>%
    writeLines()

  new_question("(section `identifiers`) Do the identifier details look complete and accurate?")
} else {
  new_question("(section `identifiers`) Are there identifiers that should be included for this dataset?")
}
```

## Taxa sampled

We have records on `r data_study$measurements$taxon_name %>% unique() %>% length()` taxa from your study. The full list of names for taxa recorded in your study is listed at the end of the report.

# Measurements

## Overview

```{r curve_overview, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# A response is a group of readings, not a table of its own: its attributes are
# columns on `measurements`, and `driver` is looked up from `data_types`.
driver_of <- function(dt) {
  vapply(dt, function(x) {
    if (is.na(x) || is.null(data_types[[x]])) return(NA_character_)
    v <- data_types[[x]][["driver"]]
    if (is.null(v) || identical(v, ".na") || is.na(v)) return(NA_character_)
    as.character(v)
  }, character(1), USE.NAMES = FALSE)
}

responses <- data_study$measurements %>%
  group_by(dataset_id, response_id) %>%
  summarise(
    data_type = data_type[[1]],
    point_order = point_order[[1]],
    taxon_name = taxon_name[[1]],
    collection_date = collection_date[[1]],
    n_points = as.character(n_distinct(point_id)),
    .groups = "drop"
  ) %>%
  mutate(driver = driver_of(data_type))
points <- response_pivot_wider(data_study, with_curves = FALSE)
value_cols <- setdiff(names(points), c("dataset_id", "response_id", "point_id"))

n_multi <- sum(as.integer(responses$n_points) > 1)

writeLines(sprintf(
  paste0("This dataset holds **%s measurement%s** on **%s taxa**, made up of %s ",
         "individual readings across %s variables.\n"),
  format(nrow(responses), big.mark = ","), ifelse(nrow(responses) == 1, "", "s"),
  length(unique(responses$taxon_name)),
  format(nrow(data_study$measurements), big.mark = ","),
  length(value_cols)
))

writeLines(sprintf(
  paste0("Of those, **%s are response curves** -- sets of readings taken across a ",
         "driver gradient, whose points belong together -- and %s are single-point ",
         "observations.\n"),
  format(n_multi, big.mark = ","),
  format(nrow(responses) - n_multi, big.mark = ",")))

responses %>%
  group_by(data_type) %>%
  summarise(
    curves = n(),
    taxa = n_distinct(taxon_name),
    points = sum(as.integer(n_points)),
    `median points per curve` = stats::median(as.integer(n_points)),
    driver = paste(unique(stats::na.omit(driver)), collapse = ", "),
    .groups = "drop"
  ) %>%
  arrange(desc(curves)) %>%
  my_kable_styling() %>%
  writeLines()
```

```{r curve_variables, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
writeLines("\nThe variables recorded, and how completely:\n")

points %>%
  select(all_of(value_cols)) %>%
  summarise(across(everything(), ~ sum(!is.na(.x)))) %>%
  tidyr::pivot_longer(everything(), names_to = "variable", values_to = "readings") %>%
  filter(readings > 0) %>%
  mutate(
    `% of points` = round(100 * readings / nrow(points), 1),
    unit = purrr::map_chr(variable, ~ definitions[[.x]]$units %||% NA_character_),
    label = purrr::map_chr(variable, ~ definitions[[.x]]$label %||% NA_character_)
  ) %>%
  arrange(desc(readings)) %>%
  my_kable_styling() %>%
  writeLines()

new_question("(section `measurements`) Are there variables you recorded that are missing from this table?")
```

```{r curve_excluded, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
real_exclusions <- subset(data_study$excluded_data, error != "Missing value")

writeLines(ifelse(nrow(real_exclusions) == 0,
  "\nAll readings passed quality control, so nothing was excluded. (Yay!)\n",
  "\nSome readings did not pass quality control and were excluded:\n"))

if (nrow(real_exclusions) > 0) {
  real_exclusions %>%
    group_by(variable, error) %>%
    summarise(`readings excluded` = n(), .groups = "drop") %>%
    arrange(desc(`readings excluded`)) %>%
    my_kable_styling() %>%
    writeLines()

  # An out-of-range reading is the commonest exclusion for instrument data, and
  # the range that rejected it is the useful thing to show alongside.
  out_of_range <- real_exclusions %>% filter(grepl("range|Range", error))

  if (nrow(out_of_range) > 0) {
    writeLines(paste0(
      "\nThese fell outside the allowed range for their variable. The ranges are ",
      "in `config/variables.yml`; if a value below is real, the range is what ",
      "needs changing.\n"))

    out_of_range %>%
      group_by(variable) %>%
      summarise(
        excluded = n(),
        `smallest excluded` = min(suppressWarnings(as.numeric(value)), na.rm = TRUE),
        `largest excluded` = max(suppressWarnings(as.numeric(value)), na.rm = TRUE),
        .groups = "drop"
      ) %>%
      mutate(
        `allowed min` = purrr::map_dbl(variable, ~ definitions[[.x]]$allowed_values_min %||% NA_real_),
        `allowed max` = purrr::map_dbl(variable, ~ definitions[[.x]]$allowed_values_max %||% NA_real_)
      ) %>%
      my_kable_styling() %>%
      writeLines()

    new_question("(section `measurements`) Are the excluded values above genuine, or a units or transcription problem? If genuine, we need to widen the allowed range.")
  }
}
```

## The curves

```{r curve_plots, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# A curve is only worth drawing if it has a driver and more than one point.
drawable <- responses %>%
  filter(!is.na(driver), as.integer(n_points) > 1)

if (nrow(drawable) == 0) {
  writeLines(paste0(
    "\nThis dataset holds no multi-point response curves, so there is nothing ",
    "to draw here. Single-point observations are summarised under ",
    "[Variables](#variables) below.\n"))
} else {
  writeLines(paste0(
    "\nEach panel is one curve: the response variable against the driver it was ",
    "measured across, with points in the order they were recorded. **This is the ",
    "check that matters most** -- a curve that looks wrong here is wrong, whatever ",
    "the summary statistics say.\n"))
}

# What to plot against the driver, per data type. Assimilation for the gas
# exchange responses, loss of conductivity for the vulnerability curves.
response_for <- function(dt, available) {
  preferred <- switch(
    dt,
    "ACi" = "A", "ACi-T" = "A", "AQ" = "A", "gs_drydown" = "gsw",
    "stem hydraulic vulnerability" = "PLCstem",
    "leaf hydraulic vulnerability" = c("PLCleaf", "kleaf"),
    "PV curve" = "RWCleaf",
    "T-F0" = "f0",
    NULL
  )
  hit <- intersect(preferred, available)
  if (length(hit) > 0) return(hit[[1]])
  NULL
}

MAX_PANELS <- 24

for (dt in sort(unique(drawable$data_type))) {

  in_type <- drawable %>% filter(data_type == dt)
  driver <- stats::na.omit(unique(in_type$driver))[[1]]
  response <- response_for(dt, value_cols)

  if (is.null(response) || !driver %in% value_cols) next

  dat <- points %>%
    filter(response_id %in% in_type$response_id) %>%
    select(all_of(c("response_id", "point_id", driver, response))) %>%
    rename(x = all_of(driver), y = all_of(response)) %>%
    mutate(across(all_of(c("x", "y")), ~ suppressWarnings(as.numeric(.x)))) %>%
    filter(!is.na(x), !is.na(y)) %>%
    left_join(in_type %>% select(all_of(c("response_id", "taxon_name"))), by = "response_id")

  if (nrow(dat) == 0) next

  shown <- utils::head(sort(unique(dat$response_id)), MAX_PANELS)
  n_total <- length(unique(dat$response_id))

  writeLines(sprintf("\n### %s\n", dt))
  writeLines(sprintf(
    "%s curve%s of this type%s. Driver: `%s`; response: `%s`.\n",
    n_total, ifelse(n_total == 1, "", "s"),
    if (n_total > MAX_PANELS) sprintf(", showing the first %d", MAX_PANELS) else "",
    driver, response))

  p <- dat %>%
    filter(response_id %in% shown) %>%
    ggplot2::ggplot(ggplot2::aes(x = x, y = y)) +
    ggplot2::geom_path(ggplot2::aes(group = response_id),
                       colour = "grey60", linewidth = 0.3) +
    ggplot2::geom_point(ggplot2::aes(colour = taxon_name), size = 1) +
    ggplot2::facet_wrap(~ response_id, scales = "free") +
    ggplot2::labs(x = driver, y = response, colour = NULL) +
    ggplot2::theme_minimal(base_size = 9) +
    ggplot2::theme(legend.position = "bottom",
                   strip.text = ggplot2::element_text(size = 7))

  rows <- ceiling(length(shown) / 4)
  writeLines(embed_plot(p, height = max(3, rows * 1.6), width = 10))
}

if (nrow(drawable) > 0) {
  new_question("(section `measurements`) Do any of the responses above look wrong -- points in the wrong order, a curve that should be two, or a shape you would not expect?")
}
```

## Curve quality

```{r curve_diagnostics, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
multi <- responses %>% filter(as.integer(n_points) > 1)

if (nrow(multi) == 0) {
  writeLines("\nNo multi-point responses in this dataset.\n")
} else {

  writeLines("\nHow long the responses are:\n")

  multi %>%
    group_by(data_type) %>%
    summarise(
      curves = n(),
      shortest = min(as.integer(n_points)),
      median = stats::median(as.integer(n_points)),
      longest = max(as.integer(n_points)),
      .groups = "drop"
    ) %>%
    my_kable_styling() %>%
    writeLines()

  # A curve of two or three points cannot be fitted, so it is worth naming.
  short <- multi %>% filter(as.integer(n_points) < 4)

  if (nrow(short) > 0) {
    writeLines(sprintf(
      paste0("\n**%s curve%s of fewer than four points.** Too short to fit a ",
             "response function to, so worth knowing whether the rest of the ",
             "curve exists somewhere.\n"),
      nrow(short), ifelse(nrow(short) == 1, " has" , "s have")))

    short %>%
      count(data_type, n_points, name = "responses") %>%
      my_kable_styling() %>%
      writeLines()

    new_question("(section `measurements`) Some responses have very few points. Are the remaining points available?")
  }

  # Point order is what lets one variable be paired with another across a
  # curve. Without it the values are present but their correspondence is an
  # assumption.
  unordered <- check_curve_pairing(data_study)

  if (nrow(unordered) > 0) {
    writeLines(sprintf(
      paste0("\n**%s curve%s no recorded point order.** The readings are all ",
             "there, but which value of one variable goes with which value of ",
             "another is not recorded -- so the curve cannot be reconstructed ",
             "reliably. This is fixed by setting `repeat_measurements_id` for ",
             "the dataset.\n"),
      nrow(unordered), ifelse(nrow(unordered) == 1, " has", "s have")))
  }

  conflicts <- check_curve_points_conflicts(data_study)

  if (nrow(conflicts) > 0) {
    writeLines(sprintf(
      paste0("\n**%s variable-point combination%s more than one value**, from ",
             "different methods. The wide view keeps the one from the lowest ",
             "`method_id`; every value is still in the long table.\n"),
      sum(!conflicts$used), ifelse(sum(!conflicts$used) == 1, " has", "s have")))

    conflicts %>%
      count(variable, method_id, name = "readings") %>%
      my_kable_styling() %>%
      writeLines()
  }
}
```

```{r driver_monotonic, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# A driver that reverses mid-curve is usually a curve that is really two, or
# points that have been sorted out of measurement order.
drawable2 <- responses %>% filter(!is.na(driver), as.integer(n_points) > 2)

if (nrow(drawable2) > 0) {
  wobbly <- purrr::map_dfr(seq_len(nrow(drawable2)), function(i) {
    row <- drawable2[i, ]
    if (!row$driver %in% value_cols) return(NULL)
    v <- points %>%
      filter(response_id == row$response_id) %>%
      arrange(point_id) %>%
      pull(row$driver) %>%
      as.numeric() %>%
      stats::na.omit()
    if (length(v) < 3) return(NULL)

    # Only substantial reversals count. On an A-Ci curve the driver is
    # *measured*, not set -- the protocol steps reference CO2 and Ci follows --
    # so it wobbles by a few umol/mol between points no matter how clean the
    # curve is. Counting every sign change flagged 73 of 171 clean curves.
    # A step smaller than 5% of the driver's range is noise, not a turn.
    span <- diff(range(v))
    if (!is.finite(span) || span == 0) return(NULL)
    d <- diff(v)
    d <- d[abs(d) >= 0.05 * span]
    if (length(d) < 2) return(NULL)
    turns <- sum(diff(sign(d)) != 0)
    if (turns <= 1) return(NULL)
    tibble::tibble(response_id = row$response_id, data_type = row$data_type,
                   taxon_name = row$taxon_name, driver = row$driver,
                   `direction changes` = turns)
  })

  if (!is.null(wobbly) && nrow(wobbly) > 0) {
    writeLines(sprintf(
      paste0("\n**The driver reverses direction more than once in %s curve%s.** ",
             "One reversal is normal -- most protocols step down and back up. ",
             "More than one, counting only steps larger than 5%% of the driver ",
             "range so measurement noise does not register, is worth a look: ",
             "it can mean two responses recorded as one, or points reordered away ",
             "from the order they were measured in. Around one curve in eight ",
             "across AusFizz trips this, so it is a prompt, not a verdict.\n"),
      nrow(wobbly), ifelse(nrow(wobbly) == 1, "", "s")))

    wobbly %>%
      arrange(desc(`direction changes`)) %>%
      utils::head(20) %>%
      my_kable_styling() %>%
      writeLines()

    new_question("(section `measurements`) In the responses above the driver reverses direction more than once. Is that the measurement protocol, or should these be separate responses?")
  }
}
```

## Variables {#variables}

```{r variable_distributions, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
writeLines(paste0(
  "\nThe distribution of every variable in this dataset, against the range ",
  "allowed for it. A value outside the range is excluded from the database, so ",
  "anything sitting near an edge is worth a look.\n"))

numeric_vars <- value_cols[purrr::map_lgl(value_cols, function(v) {
  identical(definitions[[v]]$type %||% "", "numeric")
})]

for (v in sort(numeric_vars)) {
  vals <- suppressWarnings(as.numeric(points[[v]]))
  vals <- vals[!is.na(vals)]
  if (length(vals) < 3) next

  lo <- definitions[[v]]$allowed_values_min %||% NA_real_
  hi <- definitions[[v]]$allowed_values_max %||% NA_real_

  p <- tibble::tibble(value = vals) %>%
    ggplot2::ggplot(ggplot2::aes(x = value)) +
    ggplot2::geom_histogram(bins = 40, fill = "grey40")

  if (!is.na(lo)) p <- p + ggplot2::geom_vline(xintercept = lo, colour = "firebrick", linetype = "dashed")
  if (!is.na(hi)) p <- p + ggplot2::geom_vline(xintercept = hi, colour = "firebrick", linetype = "dashed")

  p <- p +
    ggplot2::labs(
      x = sprintf("%s (%s)", v, definitions[[v]]$units %||% ""),
      y = "readings",
      subtitle = sprintf("n = %s   range in data: %s to %s   allowed: %s to %s",
                         format(length(vals), big.mark = ","),
                         signif(min(vals), 4), signif(max(vals), 4),
                         ifelse(is.na(lo), "-", lo), ifelse(is.na(hi), "-", hi))
    ) +
    ggplot2::theme_minimal(base_size = 9)

  writeLines(embed_plot(p, height = 2.2, width = 7))
}
```

## Treatments

```{r treatments_section, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
treatments <- data_study$treatments

if (is.null(treatments) || nrow(treatments) == 0) {
  writeLines("\nNo experimental treatments are recorded for this dataset.\n")
} else {

  writeLines(paste0(
    "\nWhat each treatment did, as numbers rather than as a description. This is ",
    "what lets a treatment in this study be compared with one in another -- ",
    "`elevated` means +3 &deg;C in one experiment and +4 in another.\n"))

  treatments %>%
    filter(!is.na(factor)) %>%
    select(all_of(c("treatment_context_id", "context_property", "label",
                    "factor", "value", "unit"))) %>%
    distinct() %>%
    arrange(context_property, label, factor) %>%
    my_kable_styling() %>%
    writeLines()

  unquantified <- treatments %>% filter(is.na(factor)) %>%
    distinct(context_property, label)

  if (nrow(unquantified) > 0) {
    writeLines(paste0(
      "\nThese treatment levels have no quantity recorded, because the study did ",
      "not state one:\n"))

    unquantified %>% my_kable_styling() %>% writeLines()

    new_question("(section `treatments`) Can you give a number for the treatment levels above -- how much water, how many degrees, how much CO2?")
  }
}
```

# Taxon list

We have records on the following taxa from your study. We have attempted to align taxon names with [the Australian Plant Census (APC)](https://biodiversity.org.au/nsl/services/APC) or Australian Plant Names Index (APNI), a working list of all known Australian vascular plant taxa.

## Full taxa list

The table below shows all taxa in this study, where possible using the name aligned with the APC (version = `r apcalign_data_release`). Columns are as follows:

```{r, results='asis', echo=FALSE}
schema$austraits$elements$taxa$elements %>%
  convert_list_to_df1() %>%
  my_kable_styling()
```

Clicking on the name will take you to the relevant taxonomic record in the APC or APNI, where known:

```{r, results='asis', echo=FALSE}
if (any(str_detect(data_study$taxa$taxon_name, fixed(paste0("[", dataset_id, "]"))), na.rm = TRUE)) {
  writeLines(paste0(
    "**Note:** For some taxon names in this dataset that could only be resolved to genus or family, ",
    "we have appended the dataset id to the end of the taxon name as `[dataset_id]` (e.g. `Acacia sp. [dataset_id]`). ",
    "This indicates that the name refers to a specific taxon identified only to genus (or family), rather than the entire genus, ",
    "so that records of, say, `Acacia sp.` from different datasets are not incorrectly grouped together as a single \"species\" simply because they share the same taxon name."
  ))
}
```

```{r taxon_list2, results='asis', echo=FALSE}

empty_tibble_taxa <-
  tibble::tibble(
    taxon_id = NA_character_,
    taxonomic_status = NA_character_,
    taxonomic_dataset = NA_character_,
    genus = NA_character_,
    family = NA_character_,
    binomial = NA_character_,
    trinomial = NA_character_,
    taxon_distribution = NA_character_,
    establishment_means = NA_character_
  )

empty_tibble_taxonomic_updates <-
  tibble::tibble(
    aligned_name_taxon_id = NA_character_,
    aligned_name_taxonomic_status = NA_character_
  )

tmp <- data_study$taxa %>%
  select(any_of(names(schema$austraits$elements$taxa$elements))) %>%
  bind_rows(empty_tibble_taxa)

if (nrow(tmp) < 10000) {
  tmp %>%
  mutate(
    taxon_name = ifelse(!is.na(taxon_id), as_link(taxon_id, taxon_name), taxon_name)
    ) %>%
  mutate(across(everything(), ~replace_na(.x, replace = ""))) %>%
  my_kable_styling()
} else {
  cat(crayon::red("\t\tTable too large to display\n"))
  tmp %>%
  mutate(across(everything(), ~replace_na(.x, replace = ""))) %>%
  write_csv(paste0(dataset_id, "_taxa.csv"))
}
```


```{r, results='asis', echo=FALSE}
# `updated_taxa` (before the `bind_rows(empty_tibble_taxonomic_updates)`
# placeholder-row padding below, added purely so `my_kable_styling()` always
# has something to render) reflects whether any taxonomic changes were
# actually made
taxa_columns <- names(data_study$taxa)

updated_taxa <- data_study$taxonomic_updates %>% select(-dataset_id) %>%
  filter(original_name != taxon_name)

if (nrow(updated_taxa) > 0) {
  writeLines(c(
    "To create the list of aligned taxa, we needed to make some taxonomic changes. This involved two stages.",
    "",
    "* **Stage 1**: Where possible (i.e. there was no or only a few characters difference), the name you supplied was matched automatically with a known name in APC or APNI. In some cases we may have determined that a name can only be resolved to genus. Such changes are documented in the study metadata file.  The variable `aligned_name` shows the matched name. The variable `d1` shows the number of characters difference between the `original_name` and `aligned_name`. The taxonomic status of the aligned name (i.e. a synonym, orthographic variant, etc.) is indicated in column `aligned_name_taxonomic_status`. ",
    "* **Stage 2**: If the `aligned_name` is `accepted` no additional changes were made, and the `aligned_name` is the final `taxon_name`. If the `aligned_name` has a different `aligned_name_taxonomic_status` we used the R package APCalign to update the `aligned_name` to an `accepted` taxon_name. Alternative status values are indicated in brackets to indicate if alternative uses of the name were ever applied. No updates are possible for names only documented in APNI; these are names not included in the APC.",
    "",
    "Links on `aligned_name` and `taxon_name` take you to the APC or APNI record for that name.",
    ""
  ))

  data_study$taxa %>%
    filter(taxon_name %in% updated_taxa$taxon_name) %>%
    left_join(
      by = c("taxon_name"),
      updated_taxa %>%
        bind_rows(empty_tibble_taxonomic_updates) %>%
        select(original_name, taxon_name, aligned_name, aligned_name_taxon_id, aligned_name_taxonomic_status)
    ) %>%
    bind_rows(empty_tibble_taxa) %>%
    mutate(
      d1 = purrr::map2_dbl(original_name, aligned_name, ~adist(.x, .y)) %>% as.character() %>% str_replace("0", ""),
      d2 = purrr::map2_dbl(taxon_name, aligned_name, ~adist(.x, .y)) %>% as.character() %>% str_replace("0", ""),
      aligned_name = ifelse(!is.na(aligned_name_taxon_id), as_link(aligned_name_taxon_id, aligned_name), aligned_name),
      taxon_name = ifelse(!is.na(taxon_id), as_link(taxon_id, taxon_name), taxon_name)
    ) %>%
    select(original_name, d1, aligned_name, d2, taxon_name, aligned_name_taxonomic_status) %>%
    mutate(across(everything(), ~replace_na(.x, replace = ""))) %>%
    my_kable_styling() %>%
    writeLines()
}
```

```{r taxa_list3, results='asis', echo=FALSE}
# `tmp` has an all-NA placeholder row bound on (see `taxon_list2` above) so
# the "Full taxa list" table always has something to render; filtering that
# for unknown taxa would make this section look non-empty even when every
# taxon resolved, so check the real `data_study$taxa` instead.
unknown_taxa <- data_study$taxa %>% filter(is.na(taxon_id))

if (nrow(unknown_taxa) > 0) {
  writeLines(c(
    "## Unknown taxa",
    "",
    "We were unable to find taxonomic matches for the following taxa in the data. Where genus and family are shown, the specified genus aligns with an accepted genus. The term `unplaced` appears as the `status` for names on the APNI list, but not yet reviewed by the APC.",
    ""
  ))

  unknown_taxa %>%
    select(any_of(names(schema$austraits$elements$taxa$elements))) %>%
    dplyr::select(-dplyr::any_of(c("taxon_id", "scientific_name_id", "taxon_distribution", "establishment_means"))) %>%
    mutate(across(everything(), ~replace_na(.x, replace = ""))) %>%
    my_kable_styling() %>%
    writeLines()
}
```


```{r, results='asis', echo=FALSE}
new_question("(section `taxonomic_updates`) Do these taxonomic alignments and corrections look reasonable?")
```

# Review of questions

Here is a summary of the questions we would like answers to. Clicking on the links will take you back to the appropriate spot in the document. Please copy and paste these into an email when you respond.

<font size = '3'>


```{r, results='asis', echo=FALSE}
questions() %>% responses.build:::notetaker_print_all() %>% writeLines()
```


</font>


# Session information

The technical information below may be useful for us for the sake of reproducibility, but you don't need to worry about it!

<font color='grey'>
This report was generated using the data from the commit `r util_get_SHA_link()`

And the following R environment

```{r, echo=FALSE}
sessionInfo()
```

</font>

